Adaptive Approximation for Multivariate Linear Problems with Inputs Lying in a Cone

### 2019-08-19 DingHickKritzMakRICAMChapVerD.tex
Updates from Overleaf

To be more concrete, we consider the situation where the $\lambda_{\bsk}$ are specified in terms of positive \emph{coordinate weights}, $\bsw = (w_1, \ldots, w_d)$, and positive \emph{smoothness weights}, $\bss = (s_1, s_2, \ldots)$:
\begin{equation}
    \label{DHKM:prodwts}
\lambda_{\bsk}  := \prod_{\substack{\ell =1\\ k_\ell > 0}}^d w_\ell s_{k_\ell}, \qquad \bsk \in \bbK :=\NN_0^d, \ d \in \NN.
\end{equation}
This is a generalization of our illustrative example, \eqref{DHKM:IllExWts}, in Section \ref{DHKM:secexamp} where $\bss = (1, 2^{-r}, 3^{-r}, \ldots )$.  
This form of the $\lambda_{d,\bsk}$ is considered in greater detail in Section \ref{DHKM:smoothimportance_sec}.

\begin{lemma}\label{DHKM:prodwtsupperbound}
For $\bslambda = \bigl(\lambda_{\bsk} \bigr)_{\bsk \in \bbN_0^d}$ defined in \eqref{DHKM:prodwts}
we have 
\begin{align*}
\norm[p]{\bslambda} &\le \exp\left(\frac 1p \norm[p]{\bss}^p \norm[p]{\bsw}^p  \right) \qquad \text{for general } \bss, \text{ provided that  $\norm[p]{\bss}$ is finite},\\
 \norm[p]{\bslambda} & \le \exp\left(\frac {\zeta(pr)}p \norm[p]{\bsw}^p  \right) \qquad \text{for } \bss = (1, 2^{-r}, 3^{-r}, \ldots ), \text{ provided that } pr > 1.
\end{align*}
\end{lemma}
\begin{proof}
The proof proceeds by straightforward calculation under the assumption that $\norm[p]{\bss}$ is finite:
\begin{align*}
   \norm[p]{\bslambda}^ p & 
   = \sum_{\bsk \in \NN_0^d}^\infty \lambda_{d,\bsk}^p  
   = \prod_{\ell =1}^d\left(1+ w_\ell^p \sum_{k=1}^{\infty}s_{k}^p\right) 
   = \prod_{\ell =1}^d\left(1+ w_\ell^p \norm[p]{\bss}^p \right) 
   \qquad \text{ by \eqref{DHKM:prodwts} }\\
    & = \exp\left(\sum_{\ell = 1}^d \log\bigl(1 + w_\ell^{p}\norm[p]{\bss}^p \bigr) \right)\\
 &  \le \exp\left(\sum_{\ell = 1}^d  w_\ell^{p} \norm[p]{\bss}^p \right) = \exp\left(\norm[p]{\bsw}^p\norm[p]{\bss}^p \right)
\quad \text{since }\log(1+x) \le x \ \forall x\ge 0.
\end{align*}
Taking the $p^{\text{th}}$ root completes the proof for the general case.  For the particular form of $\bss$ introduced in \eqref{DHKM:IllExWts} it follows that $\norm[p]{\bss}^p = \zeta(pr)$ for $pr > 1$.
\end{proof}

@@

Lemmas \ref{DHKM:lambdineq} and \ref{DHKM:prodwtsupperbound} imply upper bounds on $\bignorm[\rho']{\bigl(  \lambda_{\bsk_i}  \bigr)_{i = n+1}^{\infty}} = \lambda_{\bsk_{n+1}}$:
\begin{align*}
\lambda_{\bsk_{n+1}} &\le \frac{1}{n^{1/p}} \exp\left(\frac 1p \norm[p]{\bss}^p \norm[p]{\bsw}^p  \right) \qquad \text{for general } \bss, \text{ provided that  $\norm[p]{\bss}$ is finite},\\
\lambda_{\bsk_{n+1}} & \le \frac{1}{n^{1/p}} \exp\left(\frac {\zeta(pr)}p \norm[p]{\bsw}^p  \right) \  \text{for } \bss = (1, 2^{-r}, 3^{-r}, \ldots ), \text{ provided that } pr > 1.
\end{align*}
These upper bounds together with Theorems \ref{DHKM:thm_cost_non_adapt} and \ref{DHKM:PilotCostThm} imply upper bounds on the information costs of Algorithms \ref{DHKM:BallAlg} and \ref{DHKM:PilotConeAlg} for the function spaces defined in the illustrative example in Section \ref{DHKM:secexamp} for all $p > 1/r$:
\begin{equation*}
    \text{Algorithm \ref{DHKM:BallAlg}:} \quad \COST(\ALG,\calB_R,\varepsilon, R') 
    \le \left \lceil  \left(\frac{R}{\varepsilon}\right)^p  \exp\left( \zeta(pr) \norm[p]{\bsw}^p  \right)    \right \rceil,
    \end{equation*}
    \begin{multline*}
    \text{Algorithm \ref{DHKM:PilotConeAlg}:} \quad \COST(\ALG,\calC_{\bslambda,n_1,A},\varepsilon, R) 
    \\ \le \max \left( n_1, 
    \left \lceil  \left(\frac{AR}{\varepsilon}\right)^p  \exp\left( \zeta(pr) \norm[p]{\bsw}^p  \right)    \right \rceil \right).
    \end{multline*}
Note that these costs are $\calO(R^p \varepsilon^{-p})$ for all $p > 1/r$. A faster decay rate of the $\lambda_{\bsk}$ due to the faster decay rate of the smoothness coefficients implies smaller information cost.
}

>>>>

To be more concrete, we consider the situation where the $\lambda_{\bsk}$ are specified in terms of positive \emph{coordinate weights}, $\bsw = (w_1, \ldots, w_d)$, and positive \emph{smoothness weights}, $\bss = (s_1, s_2, \ldots)$:
\begin{equation}
    \label{DHKM:prodwts}
\lambda_{\bsk}  := \prod_{\substack{\ell =1\\ k_\ell > 0}}^d w_\ell s_{k_\ell}, \qquad \bsk \in \bbK :=\NN_0^d, \ d \in \NN.
\end{equation}
This is a generalization of our illustrative example, \eqref{DHKM:IllExWts}, in Section \ref{DHKM:secexamp} where $\bss = (1, 2^{-r}, 3^{-r}, \ldots )$.  
This form of the $\lambda_{\bsk}$ is considered in greater detail in Section \ref{DHKM:smoothimportance_sec}.

\begin{lemma}\label{DHKM:prodwtsupperbound}
For $\bslambda = \bigl(\lambda_{\bsk} \bigr)_{\bsk \in \bbN_0^d}$ defined in \eqref{DHKM:prodwts}
we have 
\begin{align*}
\norm[p]{\bslambda} &\le \exp\left(\frac 1p \norm[p]{\bss}^p \norm[p]{\bsw}^p  \right) \qquad \text{for general } \bss, \text{ provided that  $\norm[p]{\bss}$ is finite},\\
 \norm[p]{\bslambda} & \le \exp\left(\frac {\zeta(pr)}p \norm[p]{\bsw}^p  \right) \qquad \text{for } \bss = (1, 2^{-r}, 3^{-r}, \ldots ), \text{ provided that } pr > 1,
\end{align*}
where $\zeta$ is the Riemann zeta function.
\end{lemma}
\begin{proof}
The proof proceeds by straightforward calculation under the assumption that $\norm[p]{\bss}$ is finite:
\begin{align*}
   \norm[p]{\bslambda}^ p & 
   = \sum_{\bsk \in \NN_0^d}^\infty \lambda_{d,\bsk}^p  
   = \prod_{\ell =1}^d\left(1+ w_\ell^p \sum_{k=1}^{\infty}s_{k}^p\right) 
   = \prod_{\ell =1}^d\left(1+ w_\ell^p \norm[p]{\bss}^p \right) 
   \qquad \text{ by \eqref{DHKM:prodwts} }\\
    & = \exp\left(\sum_{\ell = 1}^d \log\bigl(1 + w_\ell^{p}\norm[p]{\bss}^p \bigr) \right)\\
 &  \le \exp\left(\sum_{\ell = 1}^d  w_\ell^{p} \norm[p]{\bss}^p \right) = \exp\left(\norm[p]{\bsw}^p\norm[p]{\bss}^p \right)
\quad \text{since }\log(1+x) \le x \ \forall x\ge 0.
\end{align*}
Taking the $p^{\text{th}}$ root completes the proof for the general case.  For the particular form of $\bss$ introduced in \eqref{DHKM:IllExWts} it follows that $\norm[p]{\bss}^p = \zeta(pr)$ for $pr > 1$.
\end{proof}

@@

Lemmas \ref{DHKM:lambdineq} and \ref{DHKM:prodwtsupperbound} imply upper bounds on $\bignorm[\rho']{\bigl(  \lambda_{\bsk_i}  \bigr)_{i = n+1}^{\infty}} = \lambda_{\bsk_{n+1}}$:
\begin{align*}
\lambda_{\bsk_{n+1}} &\le \frac{1}{n^{1/p}} \exp\left(\frac 1p \norm[p]{\bss}^p \norm[p]{\bsw}^p  \right) \qquad \text{for general } \bss, \text{ provided that  $\norm[p]{\bss}$ is finite},\\
\lambda_{\bsk_{n+1}} & \le \frac{1}{n^{1/p}} \exp\left(\frac {\zeta(pr)}p \norm[p]{\bsw}^p  \right) \  \text{for } \bss = (1, 2^{-r}, 3^{-r}, \ldots ), \text{ provided that } pr > 1.
\end{align*}
These upper bounds together with Theorems \ref{DHKM:thm_cost_non_adapt} and \ref{DHKM:PilotCostThm} imply upper bounds on the information costs of Algorithms \ref{DHKM:BallAlg} and \ref{DHKM:PilotConeAlg} for the function spaces defined in the illustrative example in Section \ref{DHKM:secexamp} for all $p > 1/r$:
\begin{equation*}
    \text{Algorithm \ref{DHKM:BallAlg}:} \quad \COST(\ALG,\calB_R,\varepsilon, R') 
    \le \left \lceil  \left(\frac{R}{\varepsilon}\right)^p  \exp\left( \zeta(pr) \norm[p]{\bsw}^p  \right)    \right \rceil,
    \end{equation*}
    \begin{multline*}
    \text{Algorithm \ref{DHKM:PilotConeAlg}:} \quad \COST(\ALG,\calC_{\bslambda,n_1,A},\varepsilon, R) 
    \\ \le \max \left( n_1, 
    \left \lceil  \left(\frac{AR}{\varepsilon}\right)^p  \exp\left( \zeta(pr) \norm[p]{\bsw}^p  \right)    \right \rceil \right).
    \end{multline*}
Note that these costs are $\calO(R^p \varepsilon^{-p})$ for all $p > 1/r$. A faster decay rate of the $\lambda_{\bsk}$ due to the faster decay rate of the smoothness coefficients, $s_k$, implies asymptotically smaller information cost.
}
